# Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the communicable diseases data
data_cmnn <- read_csv(here("data/burden-of-disease-cmnn.csv"))
```

```{r}
# Filter the dataset to include only the three countries chosen by your group.
data_cmnn_mini <- data_cmnn %>% 
  filter(Entity %in% c("South Africa", "Switzerland", "Angola"))
```

## Table of Estimates for CMNN Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries.
# You should pivot the data wider to show each country in a separate column.
data_cmnn_wide <- data_cmnn_mini %>% 
  pivot_wider(names_from = Year, values_from = `DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`)

Daly_table <- kable(data_cmnn_wide, caption = "DALY burden over time for South Africa, Switzerland and Angola")

Daly_table



# Use kable() from the knitr package to render the table.
```

## Plot Showing Trends in CMNN Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table. 

data_cmnn_mini_renamed <- data_cmnn_mini %>% 
  rename(DALYs = `DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`)


daly_plot <-  ggplot(data_cmnn_mini_renamed, aes(x = Year, y = DALYs, colour = Entity)) + geom_line() + labs(title = "DALYs over time for South Africa, Switzerland and Angola", x = " Year", y = "Disability Adjusted Life Years (DALYs)", col = "Country")

daly_plot
```

## Summary of CMNN Burden Findings

The table and the graph shows the DALY burden of Communicable Maternal, Neonatal, and Nutritional Diseases in three selected countries: Angola South Africa, and Switzerland between 1900 and 2020. According to the chart the burden of diseases remain unchanged in Switzerland, while Angola a significant decrease in the burden of disease during 1900 and 2020 was observed. In South Africa, the burden diseases increased between 1990 and 2005 and it decreased between 2006 and 2020.
