# Non-Communicable Disease Burden (NCD)

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the non-communicable diseases data
data_ncd <- read_csv(here("data/burden-of-disease-ncd.csv"))
```

```{r}
# Filter the dataset to include only the three countries chosen by your group.
data_ncd_mini <- data_ncd %>% 
  filter(Entity == c("South Africa","Switzerland","Angola"))


```

## Table of Estimates for NCD Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries. 
# You should pivot the data wider to show each country in a separate column.
data_ncd_wide <- data_ncd_mini %>%
  pivot_wider(names_from = Year, values_from = `DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`) 

kable(data_ncd_wide)

# Use kable() from the knitr package to render the table.
```

## Plot Showing Trends in NCD Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table. 
ggplot(data_ncd_mini, aes(x = Year, y = `DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`, col = Entity)) +
  geom_line() +
  theme_bw() +
  labs(x = "Year", y = "DALY (Disability-adjusted life years)", col = "Country")
```

## Summary of NCD Burden Findings

From 1990 to 2019, Angola shows consistently higher DALY rates for non-communicable diseases compared to South Africa and Switzerland, with a gradual decline over the years. South Africa experiences fluctuations in DALYs, peaking in the late 1990s before gradually declining, while Switzerland consistently has the lowest DALY burden, reflecting a relatively lower impact of non-communicable diseases. Overall, all three countries show a general downward trend in the burden of non-communicable diseases over time.
