# Overall Disease Burden

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the overall disease data
data_overall <- read_csv(here("data/burden-of-disease-all-causes.csv"))
```

```{r}
# Filter the dataset to include only the three countries chosen by your group.
data_overall_mini <- data_overall %>%
  filter(Entity %in% c("South Africa","Switzerland","Angola"))
view(data_overall_mini)
```

## Table of Estimates for Overall Disease Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries. 
# You should pivot the data wider to show each country in a separate column.
data_overall_wide <- data_overall_mini %>%
  pivot_wider(names_from = Year, values_from = `DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)`)
view(data)
# Use kable() from the knitr package to render the table.
Daly_table <- kable(data_overall_wide, caption  = "Daly burden over time for South Africa, Switzerland and Angola")
Daly_table
```

## Plot Showing Trends in Overall Disease Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table. 
data_overall_mini_short <- data_overall_mini %>%
  rename(DALYs = `DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)`)

Daly_line_graph <- ggplot(data_overall_mini_short, aes(x = Year , y = DALYs, color = Entity)) + geom_line() + labs(title = "DALYs over time for South Africa, Switzerland and Angola", col = "Country")

ggsave("Daly_line_graph.png", plot = Daly_line_graph)
```

## Summary of Overall Disease Burden Findings

Provide a brief analysis based on the data presented in the table and chart. Highlight any significant findings or patterns. About 3 sentences.
The table and chart illustrate the overall burden of disease for South Africa, Switzerland and Angola between 1990 and 2020. While Switzerland had a constant decreasing trend between 1990 and 2020, South Africa and Angola both started observing a steady decreasing trend in 2005. By 2020, the overall burden of disease for South Africa and Angola was  about six times that of Switzerland.
